Indirect Conflicts: An Exploration and Discussion of Tools, Process, and Developer Insight

### 2013-09-04 iced-2013/paper.tex
Added grounded theory and evaluation.

@@

\author{\IEEEauthorblockN{Jordan Ell}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ jell@uvic.ca}
\IEEEauthorblockN{Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ danielad@csc.uvic.ca}
}

@@

Our study was performed in two parts. First, a round of semi-structured interviews were conducted which 
addressed the larger who, what, when, where, and how about indirect conflicts. Secondly, a survey was given
out which was used to confirm what was learned from the interviews on a larger sample size as well as get
a larger opinion base on general issues.

@@

Participants were invited to participate in the interviews by email and were sent a single reminder email one week
after the initial invitation if no response had been made. We directly emailed 22 participants and ended up conducting
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use. We were quite pleased with the response rate to our initial invitation emails as
our breadth of participants was rather large spanning multiple open and closed source project and our interview 
answer saturation was quite high (to be seen in Section~\ref{sec:results}). However, we were disappointed in our word
of mouth campaign. We had asked several participants to pass along our interviews invitations to other members of
their respective organizations but did not receive any additional participants through this method.

Interview participants first answered a number of structured demographic items. Next, participants were 
asked to describe various software development processes they use or encounter in their current working
environment. These structured process questions were used for later analysis to provided context to
any answer for questions that involve indirect conflicts. They next were engaged in a
semi-structured round of questions pertaining to 12 categories of questions.

@@

Each of the categories was used as a
point of aggregation for later analysis as a way of answering the X research questions laid out in 
Section~\ref{sec:intro}. While each of the 12 question categories had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
In general, once the category was explained to the participant, little prompting was necessary as the participant
would jump right into an experience, or issue he or she had with the question or category. However, not all
participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time. We account for these non answers
in our analysis as seen in Section~\ref{sec:results}. Interviews lasted 
from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
participant had in software development as they were able to give more experiences and more informed opinions.

After the interviews, we extracted a list of answers, opinions, and experiences for each of the 12 categorical
question fields. We combined commonly mentioned opinions and answers where possible, and noted where experiences
with indirect conflicts from participants were similar but generally kept the experiences tied to the individual.
Where more direct answers were given, we used the wording of our participants rather than our own descriptive 
aggregations. The results of this preliminary analysis was a list of short hypotheses supported by direct developer
experience and opinion for the 12 categories. These hypotheses were then used as a template for creating the
second step survey, in order to further test the answers given by interview participants and validate any possible
conclusions to be drawn.

@@

We selected our survey participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two large exceptions. The software organizations
that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, Ruboss Technology Corporation, Amazon, Kano Apps, Fireworks Design, and Frost Tree Games.
However, participants who took part in the round of interviews were not asked to participate in the surveys but
rather to act as a contact point for other developers in their team, project, or organization who may be interested
in completing the survey. Aside from this aforementioned list, two large groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on their large amounts of development activity on GitHub and the Apache developers
were selected based on their software development participation on specific projects known to be used heavily
by other organizations and projects.

@@

The created survey was based off of the 12 categorical hypotheses given by the round of interviews. The survey
was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
or different opinions from those already acquired from the interviews. The survey went through two rounds of
piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
with feedback given at the end. The previous interview participants were selected based on their domain knowledge
expertise with indirect conflicts in order to provide what we think would be amount to better feedback and resulting
in a more polished survey. The final survey consisted of the following questions:
2 multiple choice questions for developer experience and demographic information; 3 level of agreement questions for an indication of
what types of environments indirect conflicts are more likely to occur;
6 level of use questions to indicate what types of software changed developers find trouble some with respect to
indirect conflicts; and 9 short answer questions for indication of when indirect conflicts occur, what types of
processes are used to prevent or react to indirect conflict, and to provide an outlet for general opinion
on the matter. Each non demographic question was made optional as it was shown 
through the interviews that some questions require more experience from participants than may be provided.

@@

After the survey was closed, preliminary analysis was applied to the short answer questions in the same manor
as the interview responses. We extracted a list of answers, opinions, and experiences for each of the 9 short
answer question fields while combining commonly mentioned answers where possible. The results again from this
preliminary analysis is a list of common and unique hypotheses to each of the 9 short answer questions.

@@

We now present our results of both the interviews and surveys conducted in regards to our research questions
outlined in Section~\ref{sec:intro}. We first present and discuss the general nature of indirect conflicts in
terms of when, what types, and how often indirect conflicts occur. We next show the findings for the process
related research questions in regards to how developers react to, processes to prevent, and reactionary steps
to indirect conflicts. Lastly, we present and discuss findings for what potential tools are wanted by developers
and what researchers are currently developing.

>>>>

\newcommand{\totalCategories}{10}

@@

\author{\IEEEauthorblockN{Jordan Ell and Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ jell@uvic.ca, danielad@csc.uvic.ca}
}

@@

Our grounded theory study was performed in two parts. First, a round of semi-structured interviews were conducted which 
addressed the larger who, what, when, where, and how about indirect conflicts. Secondly, a survey was given
out which was used to confirm what was theorized from the interviews on a larger sample size as well as obtain
larger developer opinion.

\subsection{Grounded Theory}
We approached our study based on grounded theory as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open code} to create a social theory ``grounded'' in the empirical data. For an exploratory study such as
the ours, grounded theory is well suited because involves starting from very broad and abstract type questions and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
regining the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and surveys (when open ended questions were involved) was analyzed using open coding. Open coding involves
assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
link our defined codes to categories of grounded theory such as context and consequences. The ultimate goal of grounded 
theory is to produce a theory that is tied to the empirical data collected; our final theories can be seen in
findings~\ref{itm:f1} -~\ref{itm:f9}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
3 criteria that are commonly used in evaluating grounded theory studies.

@@

Participants were invited to participate in the interviews by email and were sent a single reminder email one week
after the initial invitation if no response had been made. We directly emailed 22 participants and ended up conducting
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use. We were quite pleased with the response rate to our initial invitation emails as
our breadth of participants was rather large spanning multiple open and closed source project and our interview 
answer saturation~\cite{Guest:2006:SAP} was quite high (to be seen in Section~\ref{sec:results}).

Interview participants first answered a number of structured demographic items. Next, participants were 
asked to describe various software development experiences regarding our three main areas of focus: the nature
of indirect conflicts, current processes used regarding indirect conflicts, opinions on future and the use of
current tools. processes they use or encounter in their current working environment. These three areas of focus
were studied by having the interview participants talk about their experiences and opinions in the following
\totalCategories{} semi-structured research topics which were asked of:

@@

While each of the \totalCategories{} question categories had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
In general, once the category was explained to the participant, little prompting was necessary as the participant
would jump right into an experience, or issue he or she had with the question or category. However, not all
participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time. We account for these non answers
in our analysis and results as seen in Section~\ref{sec:results}. Interviews lasted 
from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
participant had in software development as they were able to give more experiences and more informed opinions.

%After the interviews, we extracted a list of answers, opinions, and experiences for each of the \totalCategories{} categorical
%question fields. We combined commonly mentioned opinions and answers where possible, and noted where experiences
%with indirect conflicts from participants were similar but generally kept the experiences tied to the individual.
%Where more direct answers were given, we used the wording of our participants rather than our own descriptive 
%aggregations. The results of this preliminary analysis was a list of short hypotheses supported by direct developer
%experience and opinion for the \totalCategories{} categories. These hypotheses were then used as a template for creating the
%second step survey, in order to further test the answers given by interview participants and validate any possible
%conclusions to be drawn.

@@

We selected our survey participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two large exceptions. The software organizations
that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon.
However, participants who took part in the round of interviews were not asked to participate in the surveys but
rather to act as a contact point for other developers in their team, project, or organization who may be interested
in completing the survey. Aside from this aforementioned list, two large groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development participation on specific projects known to be used heavily
utilized by other organizations and projects.

@@

The created survey was based off of the \totalCategories{} categorical hypotheses given by the round of interviews. The survey
was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
or different opinions from those already acquired from the interviews. The survey went through two rounds of
piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
with feedback given at the end. The previous interview participants were selected based on their domain knowledge
expertise with indirect conflicts in order to provide what we think would be amount to better feedback and resulting
in a more polished survey. This also allowed the previously interviewed developers to examine what hypotheses were
formed and we would be moving forward with. The final survey consisted of the following:
2 multiple choice questions for developer experience and demographic information; 3 level of agreement questions for an indication of
what types of environments indirect conflicts are more likely to occur;
6 level of use questions to indicate what types of software changed developers find trouble some with respect to
indirect conflicts; and 9 short answer questions for indication of when indirect conflicts occur, what types of
processes are used to prevent or react to indirect conflict, and to provide an outlet for general opinion
on the matter. Each non demographic question was made optional as it was shown 
through the interviews that some questions require more experience from participants than may be provided.

@@

%After the survey was closed, preliminary analysis was applied to the short answer questions in the same manor
%as the interview responses. We extracted a list of answers, opinions, and experiences for each of the 9 short
%answer question fields while combining commonly mentioned answers where possible. The results again from this
%preliminary analysis is a list of common and unique hypotheses to each of the 9 short answer questions.

\subsection{Evaluation}

Following our data collection and analysis, we re-interviewed some of our initial interview participants
in order to validate our findings. We confirmed our hypotheses as to whether or not they resonate with 
industry participants' opinions and experiences regarding indirect conflicts and as to their industrial 
applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
five participants. Those that were re-interviewed came from the range of 5-10 years of software development
experience. Re-interviewed participants were given our 8 research questions along with results and 9 findings
and asked open ended questions regarding their opinions and experiences to validate our findings. The five
participants found FILL IN HERE WHEN COMPLETE.

@@

We now present our results of both the interviews and surveys conducted in regards to our research questions
outlined in Section~\ref{sec:intro}. We first present and discuss the general nature of indirect conflicts in
terms of when, what types, and how often indirect conflicts occur. We next show the findings for the process
related research questions in regards to how developers react to, processes to prevent, and reactionary steps
to indirect conflicts.
